The 2- and 3-point surface correlation functions calculations: from novel exact continuous approach to improving methodology for discrete images

### 2023-08-21 paper.tex
Forgot apostrophe
Weather -> whether

Putting everything together, we obtain the following algorithm for calculation
of $F_{ss}$ function:
\begin{algorithmic}[1]
  \Procedure{surfsurf}{$A$}
  \State $A' \gets |H*A|$
  \Comment{$A'$ becomes an interface of $A$.}
  \State $\hat{A'} \gets F[A']$
  \Comment{Apply DFT to $A'$.}
  \State \textbf{return} $F^{-1}[ |\hat{A}|^2 / N]$
  \Comment{$N$ is a number of elements in $A$}
  \EndProcedure
\end{algorithmic}

Algorithm for surface-void function $F_{sv}$ computation is similar:
\begin{algorithmic}[1]
  \Procedure{surfvoid}{$A$}
  \State $A' \gets |H*A|$
  \Comment{$A'$ becomes an interface of $A$.}
  \State $V \gets \chi_V(A)$
  \Comment{Apply an indicator function for void phase element-wise to $A$.}
  \State $\hat{A'} \gets F[A']$
  \Comment{Apply DFT to $A'$.}
  \State $\hat{V} \gets F[V]$
  \Comment{Apply DFT to $V$.}
  \State \textbf{return} $F^{-1}[ \hat{A}\overline{\hat{V}} / N]$
  \Comment{$N$ is a number of elements in $A$}
  \EndProcedure
\end{algorithmic}

@@

\section{Computation of surface functions for sets with smooth boundary}
\label{sec:algo-precise}
Previously \cite{Samarin} we tested algorithms for digital images against known
analytic solutions. Although we obtained good results in the cases of
overlapping balls and disks, weather our algorithm performs with the same high
accuracy in general is still an open question. In this section we develop a
universal test case and novel exact algorithm for calculation of $F_{ss}$ and
$F_{sss}$ functions for two- and three-dimensional sets with smooth boundary.

@@

\subsection{A disk and a ball}
Define $f(\bm{x}) = \sum_i x_i^2$. If $\bm{x} \in \mathbb{R}^2$ when an
inequation $f(\bm{x}) \le R^2$ defines a disk of radius $R$ with the center at
the origin. In this case $F_{ss}$ depends only on distance $r$ from the origin:
\begin{equation}
  F_{ss}(r) = \left\{
  \begin{array}{ll}
    \frac{4R^2}{r\sqrt{4R^2-r^2}} & \quad r < 2R \\
    0 & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation}
The comparison of the algorithm described in \cref{sec:algo} applied to images
with different resolution against theory is shown on \cref{fig:fss-disk}.
Continuous approach results for the disk coincide with analytical $F_{ss}$ function.
\begin{figure*}[!hpt]
  \centering
  \subfigure[$3\times 3$ kernel $H$]{
    \includegraphics[width=0.45\linewidth]{images/fss-disk-3x3.png}
    \label{fig:fss-disk-3x3}}
  \hfill
  \subfigure[$7\times 7$ kernel $H'$]{
    \includegraphics[width=0.45\linewidth]{images/fss-disk-7x7.png}
    \label{fig:fss-disk-5x5}}
  \caption[]{Plot of surface-surface function for a disk with radius
    $R = 0.0334$. Resolution of the image varies from $100\times 100$ pixels to
    $3100\times 3100$.}
  \label{fig:fss-disk}
\end{figure*}

@@

\subsection{Gaussian mixture}
\label{sec:gauss}
\begin{figure*}[!hpt]
  \centering
  \subfigure[An example of a ``blob'' generated by Gaussian mixture $f(x,y)$
    which consists of four Gaussian functions. White area is where
    $f(x,y) \le 3$.] {
    \includegraphics[width=0.45\linewidth, frame]{images/blob.png}
    \label{fig:blob}}
  \hfill
  \subfigure[An example of generated ellipses.] {
    \includegraphics[width=0.45\linewidth, frame]{images/ellipses.png}
    \label{fig:ellipses}}
  \caption[]{Examples of sets with smooth boundary.}
\end{figure*}
Let us now consider more complex surfaces, including concave shapes. Take $N$
random values of each of $a_i$, $b_i$ and $\sigma_i$. Let $f$ be:
\begin{equation}
  f(x,y) = \sum_{i=1}^N \frac{1}{\sigma_i} \exp(\frac{-(x-a_i)^2-(y-b_i)^2}{\sigma_i^2})
\end{equation}
Inequality $f(x,y) \le T$ defines a ``blob'' like one which can be seen on
\cref{fig:blob}. We can use the algorithm developed in \cref{sec:fss-2d} to
compute precise values of $F_{ss}$ for this blob and compare them with values
obtained by using the algorithm in section \cref{sec:algo}. The result of this
comparison is presented on \cref{fig:fss-blob}.
\begin{figure*}[!hpt]
  \centering
  \subfigure[$3\times 3$ kernel $H$]{
    \includegraphics[width=0.45\linewidth]{images/fss-blob-3x3.png}
    \label{fig:fss-blob-3x3}}
  \hfill
  \subfigure[$7\times 7$ kernel $H'$]{
    \includegraphics[width=0.45\linewidth]{images/fss-blob-7x7.png}
    \label{fig:fss-blob-5x5}}
  \caption[]{Plot of the surface-surface function for the blob on
    \cref{fig:blob}. Values of $F_{ss}$ were calculated in semi-circle with
    radius $R = 0.3$ and angle $\phi$ varying from $0$ to $\pi$. Resolution of
    the image varies from $500\times 500$ pixels to $3000\times 3000$.}
  \label{fig:fss-blob}
\end{figure*}

\subsection{Ellipses}
\label{sec:ellipses}
Take $N$ random values $a_i$, $b_i$, $c_i$, $x^c_i$ and $y^c_i$ and define
\begin{equation}
  f_i(x, y) = a_i(x-x^c_i)^2 + b(x-x^c_i)(y-y^c_i) + c_i(y-y^c_i)^2
\end{equation}
with $a_ic_i - b_i^2 > 0$. Inequation
\begin{equation}
  \min_i f_i(x, y) < A^2
\end{equation}
defines $N$ ellipses like those which can be seen on \cref{fig:ellipses}. A
comparison between two algorithms is on \cref{fig:fss-ellipses}.
\begin{figure*}[!hpt]
  \centering
  \subfigure[$3\times 3$ kernel $H$]{
    \includegraphics[width=0.45\linewidth]{images/fss-ellipses-3x3.png}
    \label{fig:fss-ellipses-3x3}}
  \hfill
  \subfigure[$7\times 7$ kernel $H'$]{
    \includegraphics[width=0.45\linewidth]{images/fss-ellipses-7x7.png}
    \label{fig:fss-ellipses-5x5}}
  \caption[]{Plot of the surface-surface function for ellipses on
    \cref{fig:ellipses}. Values of $F_{ss}$ were calculated in semi-circle with
    radius $R = 0.1$ and angle $\phi$ varying from $0$ to $\pi$. Resolution of
    the image varies from $500\times 500$ pixels to $3000\times 3000$.}
  \label{fig:fss-ellipses}
\end{figure*}

>>>>

Putting everything together, we obtain the following algorithm for calculation
of $F_{ss}$ function:
\begin{algorithmic}[1]
  \Procedure{surfsurf}{$A$}
  \State $A' \gets |H*A|$
  \Comment{$A'$ becomes an interface of $A$.}
  \State $\hat{A'} \gets F[A']$
  \Comment{Apply DFT to $A'$.}
  \State \textbf{return} $F^{-1}[ |\hat{A'}|^2 / N]$
  \Comment{$N$ is a number of elements in $A$}
  \EndProcedure
\end{algorithmic}

Algorithm for surface-void function $F_{sv}$ computation is similar:
\begin{algorithmic}[1]
  \Procedure{surfvoid}{$A$}
  \State $A' \gets |H*A|$
  \Comment{$A'$ becomes an interface of $A$.}
  \State $V \gets \chi_V(A)$
  \Comment{Apply an indicator function for void phase element-wise to $A$.}
  \State $\hat{A'} \gets F[A']$
  \Comment{Apply DFT to $A'$.}
  \State $\hat{V} \gets F[V]$
  \Comment{Apply DFT to $V$.}
  \State \textbf{return} $F^{-1}[ \hat{A'}\overline{\hat{V}} / N]$
  \Comment{$N$ is a number of elements in $A$}
  \EndProcedure
\end{algorithmic}

@@

\section{Computation of surface functions for sets with smooth boundary}
\label{sec:algo-precise}
Previously \cite{Samarin} we tested algorithms for digital images against known
analytic solutions. Although we obtained good results in the cases of
overlapping balls and disks, whether or not our algorithm performs with the same
high accuracy in general is still an open question. In this section we develop a
universal test case and novel exact algorithm for calculation of $F_{ss}$ and
$F_{sss}$ functions for two- and three-dimensional sets with smooth boundary.

@@

\subsection{A disk and a ball}
Define $f(\bm{x}) = \sum_i x_i^2$. If $\bm{x} \in \mathbb{R}^2$ when an
inequation $f(\bm{x}) \le R^2$ defines a disk of radius $R$ with the center at
the origin. In this case $F_{ss}$ depends only on distance $r$ from the origin:
\begin{equation}
  F_{ss}(r) = \left\{
  \begin{array}{ll}
    \frac{4R^2}{r\sqrt{4R^2-r^2}} & \quad r < 2R \\
    0 & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation}
The comparison of the algorithm described in \cref{sec:algo} applied to images
with different resolution against theory is shown on \cref{fig:fss-disk}.
Continuous approach results for the disk coincide with analytical $F_{ss}$ function.
\begin{figure*}[!hpt]
  \centering
  \subfigure[$3\times 3$ kernel $H$]{
    \includegraphics[width=\linewidth]{images/fss-disk-3x3.png}
    \label{fig:fss-disk-3x3}}
  \vskip\baselineskip
  \subfigure[$7\times 7$ kernel $H'$]{
    \includegraphics[width=\linewidth]{images/fss-disk-7x7.png}
    \label{fig:fss-disk-5x5}}
  \caption[]{Plot of surface-surface function for a disk with radius
    $R = 0.0334$. Resolution of the image varies from $100\times 100$ pixels to
    $3000\times 3000$.}
  \label{fig:fss-disk}
\end{figure*}

@@

\subsection{Gaussian mixture}
\label{sec:gauss}
\begin{figure*}[!hpt]
  \centering
  \subfigure[An example of a ``blob'' generated by Gaussian mixture $f(x,y)$
    which consists of four Gaussian functions. White area is where
    $f(x,y) \le 3$.] {
    \includegraphics[width=0.45\linewidth, frame]{images/blob.png}
    \label{fig:blob}}
  \hfill
  \subfigure[An example of generated ellipses.] {
    \includegraphics[width=0.45\linewidth, frame]{images/ellipses.png}
    \label{fig:ellipses}}
  \caption[]{Examples of sets with smooth boundary.}
\end{figure*}
Let us now consider more complex surfaces, including concave shapes. Take $N$
random values of each of $a_i$, $b_i$ and $\sigma_i$. Let $f$ be:
\begin{equation}
  f(x,y) = \sum_{i=1}^N \frac{1}{\sigma_i} \exp(\frac{-(x-a_i)^2-(y-b_i)^2}{\sigma_i^2})
\end{equation}
Inequality $f(x,y) \le T$ defines a ``blob'' like one which can be seen on
\cref{fig:blob}. We can use the algorithm developed in \cref{sec:fss-2d} to
compute precise values of $F_{ss}$ for this blob and compare them with values
obtained by using the algorithm in section \cref{sec:algo}. The result of this
comparison is presented on \cref{fig:fss-blob}.
\begin{figure*}[!hpt]
  \centering
  \subfigure[$3\times 3$ kernel $H$]{
    \includegraphics[width=\linewidth]{images/fss-blob-3x3.png}
    \label{fig:fss-blob-3x3}}
  \vskip\baselineskip
  \subfigure[$7\times 7$ kernel $H'$]{
    \includegraphics[width=\linewidth]{images/fss-blob-7x7.png}
    \label{fig:fss-blob-5x5}}
  \caption[]{Plot of the surface-surface function for the blob on
    \cref{fig:blob}. Values of $F_{ss}$ were calculated in semi-circle with
    radius $R = 0.3$ and angle $\phi$ varying from $0$ to $\pi$. Resolution of
    the image varies from $500\times 500$ pixels to $3000\times 3000$.}
  \label{fig:fss-blob}
\end{figure*}

\subsection{Ellipses}
\label{sec:ellipses}
Take $N$ random values $a_i$, $b_i$, $c_i$, $x^c_i$ and $y^c_i$ and define
\begin{equation}
  f_i(x, y) = a_i(x-x^c_i)^2 + b(x-x^c_i)(y-y^c_i) + c_i(y-y^c_i)^2
\end{equation}
with $a_ic_i - b_i^2 > 0$. Inequation
\begin{equation}
  \min_i f_i(x, y) < A^2
\end{equation}
defines $N$ ellipses like those which can be seen on \cref{fig:ellipses}. A
comparison between two algorithms is on \cref{fig:fss-ellipses}.
\begin{figure*}[!hpt]
  \centering
  \subfigure[$3\times 3$ kernel $H$]{
    \includegraphics[width=\linewidth]{images/fss-ellipses-3x3.png}
    \label{fig:fss-ellipses-3x3}}
  \vskip\baselineskip
  \subfigure[$7\times 7$ kernel $H'$]{
    \includegraphics[width=\linewidth]{images/fss-ellipses-7x7.png}
    \label{fig:fss-ellipses-5x5}}
  \caption[]{Plot of the surface-surface function for ellipses on
    \cref{fig:ellipses}. Values of $F_{ss}$ were calculated in semi-circle with
    radius $R = 0.1$ and angle $\phi$ varying from $0$ to $\pi$. Resolution of
    the image varies from $500\times 500$ pixels to $3000\times 3000$.}
  \label{fig:fss-ellipses}
\end{figure*}
